---
title: "gsea"
output: html_document
params:
  analysis_name: GSEA_test
  working_dir: ~/projects/Jeremy_Chan/in_class_exercises/gsea/data/
  output_dir: ~/projects/Jeremy_Chan/in_class_exercises/gsea/output/
  rnk_file: MesenvsImmuno_RNASeq_ranks.rnk
  classes_file: classes.cls
  expression_file: expression.txt
  run_gsea: yes
  gsea_jar: ~/projects/Jeremy_Chan/GSEA_4.4.0/gsea-cli.sh
  gsea_directory: ''
  fdr_thresh: 0.01
  pval_thresh: 1
  java_version: 11
  
---

I installed the GSEA software from their website (command-line version) and
unzipped it into my directory. I put the path in the parameters of the
Rmd file (see above) - the parameters section was modified from the enrichment
map analysis Rmd.

Set parameters first:
```{r}
gsea_jar <- params$gsea_jar
working_dir <- params$working_dir
output_dir <- params$output_dir
analysis_name <- params$analysis_name
rnk_file <- params$rnk_file
run_gsea <- params$run_gsea
dest_gmt_file = ""
```

Get the .gmt file:
- Had to change http to https (contents showed that the web address was moved)
- Homework assignment states to use the September 1, 2025 dataset so I changed current_release to September_01_2025
- We want gene sets with GO biological process, no IEA and pathways
- I think it's 'Human_GOBP_AllPathways_noPFOCR_no_GO_iea_September_01_2025_symbol.gmt'
- There's also with PFOCR (pathway figure OCR) but I will use the one without for this test
```{r}
if(dest_gmt_file == "") {
  gmt_url = "https://download.baderlab.org/EM_Genesets/September_01_2025/Human/symbol/"
  
  filenames = getURL(gmt_url)
  tc = textConnection(filenames)
  contents = readLines(tc)
  close(tc)
  
  rx = gregexpr("(?<=<a href=\")(.*.GOBP_AllPathways_noPFOCR_no_GO_iea.*.)(.gmt)(?=\">)",
    contents, perl = TRUE)
  gmt_file = unlist(regmatches(contents, rx))
  dest_gmt_file <- file.path(output_dir, gmt_file)
  
  if(!file.exists(dest_gmt_file)){
    download.file(
      paste(gmt_url,gmt_file, sep=""),
      destfile = dest_gmt_file
    )
  }
}

```
Successfully downloaded the .gmt file into the output directory.

Now, we want to run GSEA on these parameters:
- maximum geneset size of 200
- minimum geneset size of 15
- gene set permutation

Had to change permission for the jar file (gsea-cli.sh) for it to run properly ('chmod +x GSEA_4.4.0/gsea-cli.sh').
```{r}
if(run_gsea) {
  command <- paste("", gsea_jar,
                   "GSEAPreRanked -gmx", dest_gmt_file,
                   "-rnk", file.path(working_dir, rnk_file),
                   "-collapse false -nperm 1000 -scoring_scheme weighted",
                   "-rpt_label ", analysis_name,
                   "-plot_top_x 20 -rnd_seed 12345 -set_max 200",
                   "-set_min 15 -zip_report false",
                   "-out ", output_dir,
                   " > gsea_output.txt", sep = " ")
  system(command)
}
```

We can see an overview of the information in the created 'index.html' file!